---
title: "R Notebook"
author: "Eduardo Rojo"
date: "May 4th, 2023"
output:
  html_document:
    df_print: paged
    toc: true
    toc_depth: 4
    toc_float: true
---

# This is the header of the document {#markdown-header}
\
\
\
```{r download-data, echo = FALSE, include = FALSE, message = FALSE, warning = FALSE}
#### First plot
dir.create("data")
download.file(url="https://raw.githubusercontent.com/cambiotraining/reproducibility-training/master/data/gapminder_data.csv", destfile="data/gapminder_data.csv")
```

```{r load-library, message = FALSE, warning = FALSE, echo = FALSE}
#load tidyverse library
library(tidyverse) # used for data manipulation
library(rmarkdown) # used for paged_table function
library(kableExtra) # used for table
library(ggpubr) #used for ggarrange function 

#read file into R
pop_data <- read_csv("data/gapminder_data.csv")

#create a table with data from European countries in 2007 showing the countries 
#with the largest life expectancy at the top 
euro_data_tbl <- pop_data %>% 
                    filter(continent == "Europe" & year == 2007) %>% 
                    select(-continent, -year) %>% 
                    arrange(desc(lifeExp)) %>% 
                    rename(Country = country, 
                           "Population Size" = pop, 
                           "Life Expectancy" = lifeExp, 
                           "GDP" = gdpPercap)
    
kbl(euro_data_tbl) %>%
          kable_styling(bootstrap_options = c("striped", "hover"))

```

```{r paged-table}
paged_table(euro_data_tbl)

```


Next, the life expectancy in Europe is observed across different years.  The aim of this is to check if there was a change in life expectancy over the years.  The data used for the life expectancy over the years analysis is the gapminder data and the following data manipulation was performed:

```{r fig-data}
#keep on European data and change year to factor
euro_data_fig <- pop_data %>% 
                    filter(continent == "Europe")  %>% 
                    mutate(year=as_factor(year))

```

The life expectancy for Europe over the years is plotted as following:

```{r fig, warning=FALSE, fig.cap="Life Expectancy in Europe over the years"}
#Euro plot
euro_plot <- euro_data_fig %>%
                ggplot(mapping=aes(x=year, y=lifeExp)) +
                geom_violin() +
                stat_summary(fun.y = median)

#draw euro plot
euro_plot

```


Comparison of life expectancy from Europe and United Kingdom:
```{r fig-comp, warning=FALSE, fig.cap="Life Expectancy in European and UK over the years"}
#keep only United Kingdom data and change year to factor
uk_data_fig <- pop_data %>%
                    filter(country == "United Kingdom") %>%
                    mutate(year=as_factor(year))

#UK plot
uk_plot <- uk_data_fig %>%
                ggplot(mapping=aes(x=year, y=lifeExp)) +
                geom_point()

#draw euro plot next to UK plot
ggarrange(euro_plot, uk_plot, ncol=2, nrow=1, labels=c("A","B"))
```


### Lists
#### Indexed lists
1. Item 1
2. Item 2
3. Item 3

#### Indexed lists with spacing
1. Item 1
2. Item 2
3. Item 3
    a. Item 3a
        i.  Item 3ai
        ii. Item 3aii
    b. Item 3b
    c. Item 3c
4. Item 4

#### Lists with vignettes
* Item 1
* Item 2
* Item 3
    * Item 3a
        *  Item 3ai
        * Item 3aii
    * Item 3b
    * Item 3c
* Item 4

#### To-do lists
- [ ] an unchecked task list item
- [x] checked item


### Tables

1. My first table

| Column 1    | Column 2    | Column 3    |
| ----------- | ----------- | ----------- |
| Item 1,1    | Item 1,2    | Item 4,5    |
| Item 2,1    | Item 2,2    | Item 5,6    |


2. My first centered table

| Column 1    | Column 2    | Column 3    |
|    :----:   |    :----:   |    :----:   |
| Item 1,1    | Item 1,2    | Item 1,3    |
| Item 2,1    | Item 2,2    | Item 2,3    |


### Blocks
> Testing the block with a "quote"


[Go to header](#markdown-header) \
\
\
[Go to training website](https://cambiotraining.github.io/reproducibility-training/rmarkdown.html)
[kable styling](https://cran.r-project.org/web/packages/kableExtra/vignettes/awesome_table_in_html.html#Table_Styles)
[Examples of R notebooks: rpubs](https://rpubs.com/)